---
title: "Donald Trump  Twitter Analysis"
author: "Ryan Buckner, Raiyan Kabir, Mariam Ghavalyan, Chinmaya Holla"
---

###Introduction


Texty
Text
Text  

&nbsp;

###Relationships Between Words
TextTextTextTextTextTextTextText

```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.align='center'}
setwd("/Users/Ryan/Desktop")

library(twitteR)
library(SnowballC)
library(wordcloud)
library(tm)
library(stringr)
library(dplyr)
library(tidytext)
library(janeaustenr)
library(twitteR)
library(stringr)
library(tidyr)
library(igraph)
library(ggraph)
library(devtools)
library(widyr)
library(topicmodels)
library(ggplot2)
library(syuzhet)
library(lubridate)
library(scales)


#######################################################################################################################
############  CLEAN DATA 
#################################################################################################################

#Read csv
hate <- read.csv("all_hate_tweets_FINAL.csv")


#Clean data
hate$value <- gsub("(RT|via)((?:\\b\\W*@\\w+)+)", "", hate$value)
hate$value <- gsub("http\\w+", "", hate$value)
hate$value <- gsub("@\\w+", "", hate$value)
hate$value <- gsub("@", "", hate$value)
hate$value <- gsub("[[:punct:]]", "", hate$value)
hate$value <- gsub("[[:digit:]]", "", hate$value)
hate$value <- gsub("&amp", "", hate$value)
hate$value <- gsub("[ \t]{2,}", "", hate$value)
hate$value <- gsub("^\\s+|\\s+$", "", hate$value) 
hate$value <- gsub("tco.*? ", "", hate$value) 
hate$value <- gsub("tco.*", "", hate$value)
hate$value <- gsub("http.*? ", "", hate$value) 
hate$value <- gsub("amp", "", hate$value)
hate$value <- gsub("http", "", hate$value)
hate <- hate[!(is.na(hate$value) | hate$value==""), ]

#######################################################################################
#### FIRST STUFF
#######################################################################################


trump1 <- read.csv("TruthFacts.csv", stringsAsFactors=FALSE)

trump1$text <- gsub("(RT|via)((?:\\b\\W*@\\w+)+)", "", trump1$text)
trump1$text <- gsub("http\\w+", "", trump1$text)
trump1$text <- gsub("@\\w+", "", trump1$text)
trump1$text <- gsub("@", "", trump1$text)
trump1$text <- gsub("[[:punct:]]", "", trump1$text)
trump1$text <- gsub("[[:digit:]]", "", trump1$text)
trump1$text <- gsub("&amp", "", trump1$text)
trump1$text <- gsub("[ \t]{2,}", "", trump1$text)
trump1$text <- gsub("^\\s+|\\s+$", "", trump1$text) 
trump1$text <- gsub("tco.*? ", "", trump1$text)
trump1$text <- gsub("\"", "", trump1$text)
trump1$text <- gsub("tco.*", "", trump1$text) 
trump1$text <- gsub("MakeAmericaGreatAgain", "", trump1$text)


#######################################################################################
#### NGRAMS
#######################################################################################


trump_bigrams <- trump1 %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2)

#####
# COUNT AND FILTERING N-GRAMS

#trump_bigrams %>%
#  count(bigram, sort = TRUE)

#STOP-WORDS

bigrams_separated <- trump_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

# new bigram counts:
bigram_counts <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)

#recombined words
bigrams_united <- bigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ")


#######################################################################################
#### VISUALIZING A NETWORK OF BIGRAMS WITH GGRAPH
#######################################################################################


# filter for only relatively common combinations
bigram_graph <- bigram_counts %>%
  filter(n > 20) %>%
  graph_from_data_frame()

```
Text
Text
Text
```{r 2015-12-13_sentiment_graph, echo=FALSE, fig.width=10, fig.height=7}
set.seed(2016)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                 arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()
```
TEXT ABOUT GRAPH ABOVE


&nbsp;


###Pairwise Correlation
texttexttexttexttexttexttext

```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.align='center'}
#######################################################################################
#### COUNTING AND CORRELATION AMONG SECTIONS
#######################################################################################

trump_section_words <- trump1 %>%
  mutate(section = row_number() %/% 10) %>%
  filter(section > 0) %>%
  unnest_tokens(word, text) %>%
  filter(!word %in% stop_words$word)


# count words co-occuring within sections
word_pairs <- trump_section_words %>%
  pairwise_count(word, section, sort = TRUE)

#######################################################################################
#### PAIRWISE CORRELATION
#######################################################################################


# we need to filter for at least relatively common words first
word_cors <- trump_section_words %>%
  group_by(word) %>%
  filter(n() >= 20) %>%
  pairwise_cor(word, section, sort = TRUE)

```



- Religious groups
```{r pairwise_correlation_graph, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}
word_cors %>%
  filter(item1 %in% c("christian", "jewish", "muslim", "athiest", "god")) %>%
  group_by(item1) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(item2 = reorder(item2, correlation)) %>%
  ggplot(aes(item2, correlation, fill=correlation)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ item1, scales = "free") +
  coord_flip()
```
graph description

&nbsp;


- Politicians
```{r pairwise_correlation_two_graph, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}
word_cors %>%
  filter(item1 %in% c("hillary", "jeb", "ted", "bernie")) %>%
  group_by(item1) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(item2 = reorder(item2, correlation)) %>%
  ggplot(aes(item2, correlation, fill=correlation)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ item1, scales = "free") +
  coord_flip()
```
graph description

&nbsp;


- Policies
```{r pairwise_correlation_three_graph, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}
word_cors %>%
  filter(item1 %in% c("healthcare", "economy", "immigration", "war", "terror")) %>%
  group_by(item1) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(item2 = reorder(item2, correlation)) %>%
  ggplot(aes(item2, correlation, fill=correlation)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ item1, scales = "free") +
  coord_flip()
```
graph description

&nbsp;

- Race
```{r pairwise_correlation_four_graph, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}
word_cors %>%
  filter(item1 %in% c("white", "black", "hispanic", "arab")) %>%
  group_by(item1) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(item2 = reorder(item2, correlation)) %>%
  ggplot(aes(item2, correlation, fill=correlation)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ item1, scales = "free") +
  coord_flip()
```
graph description

&nbsp;

###Sentiment Analysis
TextTextTextText
&nbsp;
```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.align='center'}
#######################################################################################
#######################################################################################
#######################################################################################
#### SENTIMENT ANALYSIS
#######################################################################################
#######################################################################################
#######################################################################################

#get_sentiments("nrc")

#mySentiment <- get_nrc_sentiment(trump1$text)
#######################################################################################
#### SENTIMENT ANALYSIS WITH INNER JOIN
#######################################################################################


mySentiment <- get_nrc_sentiment(trump1$text)

tweets <- cbind(trump1, mySentiment)
```
- Total sentiment score for all tweets
```{r sentiment_all_tweets_graph, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}
sentimentTotals <- data.frame(colSums(tweets[,c(7:16)]))
names(sentimentTotals) <- "count"
sentimentTotals <- cbind("sentiment" = rownames(sentimentTotals), sentimentTotals)
rownames(sentimentTotals) <- NULL
ggplot(data = sentimentTotals, aes(x = sentiment, y = count)) +
  geom_bar(aes(fill = sentiment), stat = "identity") +
  theme(legend.position = "none") +
  xlab("Sentiment") + ylab("Total Count") + ggtitle("Total Sentiment Score for All Tweets")
```
description of graph above

&nbsp;


```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.align='center'}
tidy_trump <- trump1 %>%
  group_by(year) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

nrcjoy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

#tidy_books %>%
 # filter(year == "2017") %>%
 # inner_join(nrcjoy) %>%
 # count(word, sort = TRUE)

trumpsentiment <- tidy_trump %>%
  inner_join(get_sentiments("nrc")) %>%
  count(year, index = linenumber %/% 1, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```


- Sentiment per tweet
```{r sentiment_timeline_graph, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}
ggplot(trumpsentiment, aes(index, sentiment, fill = year)) +
  geom_col(show.legend = FALSE) 
```
Explanation of graph

&nbsp;

###Topic Modeling
Text Text Text
&nbsp;
```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.align='center'}
#######################################################################################
#######################################################################################
#######################################################################################
#### TOPIC MODELING
#######################################################################################
#######################################################################################
#######################################################################################

# divide into documents, each representing one chapter
by_handle <- hate %>%
  group_by(group_id) %>%
  mutate(chapter = cumsum(str_detect(value, regex("^asdf ", ignore_case = TRUE)))) %>%
  ungroup() %>%
  filter(chapter > 0) %>%
  unite(document, group_id, chapter)

# split into words
by_handle_word <- by_handle %>%
  unnest_tokens(word, value)

# find document-word counts
word_counts <- by_handle_word %>%
  anti_join(stop_words) %>%
  count(document, word, sort = TRUE) %>%
  ungroup()


#######################################################################################
#### LDA ON CHAPTERS
#######################################################################################

handle_dtm <- word_counts %>%
  cast_dtm(document, word, n)


handle_lda <- LDA(handle_dtm, k = 11, control = list(seed = 1234))


handle_topics <- tidy(handle_lda, matrix = "beta")


top_terms <- handle_topics %>%
  group_by(topic) %>%
  top_n(5, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)
```

- LDA by ideological group
```{r topic_graph, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}
top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()
```
Explanation of graph

&nbsp;

###Per Document Classification
Text text text text
```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.align='center'}
#######################################################################################
#### PER DOCUMENT CLASSIFICATION
#######################################################################################

handle_gamma <- tidy(handle_lda, matrix = "gamma")


handle_gamma <- handle_gamma %>%
  separate(document, c("group_id", "chapter"), sep = "_", convert = TRUE)


handle_classifications <- handle_gamma %>%
  group_by(group_id, chapter) %>%
  top_n(1, gamma) %>%
  ungroup()

handle_classifications

handle_topics <- handle_classifications %>%
  count(group_id, topic) %>%
  group_by(group_id) %>%
  top_n(1, n) %>%
  ungroup() %>%
  transmute(consensus = group_id, topic)

handle_classifications %>%
  inner_join(handle_topics, by = "topic") %>%
  filter(group_id != consensus)

#######################################################################################
#### BY WORD ASSIGNMENTS: AUGMENT
#######################################################################################
handle_lda <- LDA(handle_dtm, k = 5, control = list(seed = 1234))

assignments <- augment(handle_lda, data = handle_dtm)

assignments <- assignments %>%
  separate(document, c("title", "chapter"), sep = "_", convert = TRUE) %>%
  inner_join(handle_topics, by = c(".topic" = "topic"))

```
- Confusion matrix
```{r assignments_graph, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}
assignments %>%
  count(title, consensus, wt = count) %>%
  group_by(title) %>%
  mutate(percent = n / sum(n)) %>%
  ggplot(aes(consensus, title, fill = percent)) +
  geom_tile() +
  scale_fill_gradient2(high = "red") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank()) +
  labs(x = "Ideological group tweets were assigned to",
       y = "Ideologial group tweets came from",
       fill = "% of assignments")
```
Explanation of graph

&nbsp;

```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.align='center'}
wrong_words <- assignments %>%
  filter(title != consensus)
```
- Wrong words
```{r wrong_words_graph, echo=FALSE, message=FALSE, fig.width=10, fig.height=7}
wrong_words %>%
  count(title, consensus, term, wt = count) %>%
  ungroup() %>%
  arrange(desc(n))
```
  
###Conclusion
  
  